### 2025-08-08 manuscript/manuscript.qmd
add another paragraph to introduction

>>>>

Research across several scientific disciplines that show that women tend to be less represented in authors compared to men (Banks et al., 2025; @son2022scientific]. Encouragingly, however, there is a general trend that the ratio of women to men in authorships is improving. From around 1960 to 2021, studies show a general improvement [@gonzalez2020women]; Ioannadis et al., 2023; Jemielniak & Wilamowski, 2025). However, there is evidence that this improvement is plateauing (Jemielniak & Wilamowski, 2025) and that although women are more likely to be first author, women are still generally unfairly underrepresented as the last author [@gonzalez2020women; @rock2021quantifying; @shah2021gender]. There could be several reasons for such a disparity: despite efforts to give fair credit for author contributions [@ni2021gendered], for example, using the Contributor Roles Taxonomy (CRediT) system, assigning authorships can be unclear. What counts as an authorship may also differ between academic disciplines. Women are more likely to experience authorship disagreements (Ni et al., 2021). There are general disagreements of when to decide authorship - women preferring to discuss in earlier stages, but men choosing authorship at the final stage. Another reason could also come from parenthood and parental leave, which account for about ~40% of gender gap in career advancement (Neilsen et al 2023). Importantly, there are strides to improving this, with opinion papers discussing way to improve receipt and reporting of intellectual credit (Banks et al., 2025) as well as creating equitable environments in academic science (Martinez Acosta et al., 2025).
